---
title: "4th_corner_test2"
author: "Tessa R Smith"
date: "2024-06-30"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
#STEP 0.1.1: Change the location for your computer. 
# You can safely use either "\\" or "/" in the path.
knitr::opts_chunk$set(echo = TRUE, root.dir = "C:/Users/Tessas/OneDrive - University ofTasmaniaTessa_PhD/WORK_Code/CH2_LOCALSCALE/CH2_4thCornerAnalysis_20241114")
```

######################## 

## INTRODUCTION

######################## 

**BEETLE BIOGEOGRAPHY PROJECT**, University of Tasmania. Data collected
by Tessa R Smith and volunteers (2020-2022). Code by Tessa Smith based
on Matthew Fielding Bass Islands birds project, last edited December
2024.

# **CODE FOR TESSA SMITH THESIS CHAPTER 2**

**AIMS**

-   **AIM 1: 4th corner analysis by traits**

-   **AIM 2: 4th corner analysis by species**

-   **AIM 3: Estimation of beetle biomass per site**

######################## 

## DATA PREPARATION

######################## 

```{r}
#STEP 0.1.2: Load packages
library(ade4)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(ncf)
library(gridExtra)
library(spdep)
library(sf)
library(tidyr)
library(readxl)
library(mvabund)

#packages deleted- check it runs again
#scales, grid, ggspatial, performance, splines, MASS, emmeans
```

```{r}
#STEP 0.2.1: Import data (environmental data)
#Reference: Own data (Tessa Smith PhD)
#File type: CSV #Format: wide form #Columns: environmental variables #Rows:Sites
ENVIRONMENTDATA <- read.csv("TS_Site_Enviro_metadata_20230622_noplants.csv", header = TRUE, stringsAsFactors = FALSE)


#Remove columns with missing data
ENVIRONMENTDATA <- ENVIRONMENTDATA[, !names(ENVIRONMENTDATA) %in% c("volume_L", "in_log_covp", "out_plant_covp", "out_moss_covp", "out_barg_covp", "out_rock_covp", "p_Beyeria", "p_Exocarpus", "p_Arthrotaxis", "p_Bursaria", "p_Cassinia", "times_burned", "time_since_fire", "logs_dist", "BIO11", "BIO4")]

# List of site_IDs to be removed
site_ids_to_remove <- c("CORINNA002", "MOUNTTFIELD001", "MTDUNADAS001", 
                        "STRATHGORDON001", "STREZLECKI001", "STREZLECKI003", 
                        "TIMBSTRACK001", "WALDHEIM01")

# Remove rows from ENVIRONMENTDATA where site_ID is in the list
ENVIRONMENTDATA <- ENVIRONMENTDATA[!ENVIRONMENTDATA$site_ID %in% site_ids_to_remove, ]

# Optionally, check the result
print(ENVIRONMENTDATA)

View(ENVIRONMENTDATA) #Opens data in another tab so it can be easily viewed
```

Notes: Keeping p_Dianella which correlates with p_Beyeria, p_Exocarpus

```{r}
#STEP 0.2.2: Import data (species records)
#Reference: Own data (Tessa Smith PhD)
#File type: CSV #Format: long form. #Columns: site_ID, species, specimen_count. #Rows: Individual beetle records
beetle_records.df <- read.csv("TS_Site_BeetleRecords_20240306.csv", header = TRUE, stringsAsFactors = FALSE)

#Reshapes species list of beetle records from long form into site.count form (wide)
#Only necessary once before start of analysis
df_reshaped <- beetle_records.df %>%
#Change from long to wide format
pivot_wider(id_cols = site_ID,
            names_from = species,
            values_from = species_count,
            # Sum the entries where there is more than 1 value in specimen_count
            values_fn = list(species_count= sum)) 

#Change NAs in the new dataset to 0
df_reshaped[is.na(df_reshaped)] <-0 
df_reshaped <- na.omit(df_reshaped)

#Change the name
INSECTDATA <- df_reshaped
view (INSECTDATA) #Opens data in another tab so it can be easily viewed
```

```{r}
#STEP 0.2.3: Remove columns from INSECTDATA that we don't want in the database

#create list of names of columns that we want to remove

columns_to_remove <- c("UNKNOWN_SPECIES",
"UNKNOWN_SPECIES-Pieces",
"PELORIDIIDAE_indet.",
"Hemiodoecus_leai",
"Hemiodoecellus_fidelis",
"Choerocoris_paganus",
"LYCIDAE_indet._(larvae)",
"Misc._beetles",
"Hemiptera_(misc.)",
"Spiders_(misc.)",
"Opiliones_(misc.)",
"Mites_(misc.)",
"Collembola_(misc.)",
"Millipedes_&_entipedes (misc.)",
"Earwigs_(misc.)",
"Hymenoptera_(misc.)",
"Snails_(misc.)",
"Other_non-beetle_invertebrates_(misc.)", 
"STAPHYLINIDAE_indet.")

#STEP 0.3.5: remove the columns
# Remove the columns from the dataframe
INSECTDATA <- INSECTDATA[ , !(names(INSECTDATA) %in% columns_to_remove)]

# View the updated dataframe
view(INSECTDATA)
```

```{r}
#STEP 0.2.4.1: Remove rare species (species present in <3 sites)
#ALTERNATIVE TO REMOVE FROM <5 SITES BELOW, choose the one you want

# Function to remove columns where species is present in less than 3 sites
remove_infrequent_species <- function(INSECTDATA, threshold = 3) {
  # Calculate the number of sites each species is present in
  species_counts <- colSums(INSECTDATA != 0)
  
  # Identify columns to keep and to remove
  columns_to_keep <- species_counts >= threshold
  columns_to_remove <- species_counts < threshold
  
  # Filter the data frame
  INSECTDATA_filtered <- INSECTDATA[, columns_to_keep]
  
  # Print the number of columns included and removed
  num_included <- sum(columns_to_keep)
  num_removed <- sum(columns_to_remove)
  cat("Number of columns included:", num_included, "\n")
  cat("Number of columns removed:", num_removed, "\n")
  
  return(INSECTDATA_filtered)
}

# Apply the function to INSECTDATA
INSECTDATA_filtered <- remove_infrequent_species(INSECTDATA, 3)

# Remove rows from INSECTDATA_filtered where site_ID is not in ENVIRONMENTDATA
INSECTDATA_filtered <- INSECTDATA_filtered[INSECTDATA_filtered$site_ID %in% ENVIRONMENTDATA$site_ID, ]

# Change name back to original name
INSECTDATA <- INSECTDATA_filtered

# View the result
view(INSECTDATA)
```

Notes: works. Final number of columns included is 226. Does not contain
NAs

```{r}
#STEP 0.2.4.2: Remove rare species (species present in <5 sites)
# Function to remove columns where species is present in less than 5 sites
remove_infrequent_species <- function(INSECTDATA, threshold = 5) {
  # Calculate the number of sites each species is present in
  species_counts <- colSums(INSECTDATA != 0)
  
  # Identify columns to keep and to remove
  columns_to_keep <- species_counts >= threshold
  columns_to_remove <- species_counts < threshold
  
  # Filter the data frame
  INSECTDATA_filtered <- INSECTDATA[, columns_to_keep]
  
  # Print the number of columns included and removed
  num_included <- sum(columns_to_keep)
  num_removed <- sum(columns_to_remove)
  cat("Number of columns included:", num_included, "\n")
  cat("Number of columns removed:", num_removed, "\n")
  
  return(INSECTDATA_filtered)
}

# Apply the function to INSECTDATA
INSECTDATA_filtered <- remove_infrequent_species(INSECTDATA, 3)

# Remove rows from INSECTDATA_filtered where site_ID is not in ENVIRONMENTDATA
INSECTDATA_filtered <- INSECTDATA_filtered[INSECTDATA_filtered$site_ID %in% ENVIRONMENTDATA$site_ID, ]

# Change name back to original name
INSECTDATA <- INSECTDATA_filtered

# View the result
view(INSECTDATA)
```

```{r, echo = FALSE}
#STEP 0.3.1: Import data (beetle trait data)
#Reference: Own data (Tessa Smith PhD)
#File type: CSV
TRAITDATA <- read.csv("TS_Site_TraitRecords_20230628.csv", header = TRUE, stringsAsFactors = FALSE)

#Remove columns with missing data
TRAITDATA <- TRAITDATA[, !names(TRAITDATA) %in% c("total.count", "no.sites", "prop.site", "vagility", "logs_dist")]

# Remove rows with NA in the specified columns
TRAITDATA <- TRAITDATA[complete.cases(TRAITDATA[, c("diet", "size", "native", "size.1")]), ]

View(TRAITDATA) #Opens data in another tab so it can be easily viewed
```

#Notes: Matt Fielding #b.df = site.count (sites as rows, species as
columns) = INSECTDATA #t.df = trat.data (bird species as rows, species
traits as columns) = TRAITDATA #s.df = site.env (sites as rows, site
traits ascolumns) = ENVIRONMENTDATA

```{r}
#STEP 0.4.1: Convert TRAITDATA into a data frame with species_ID as row names
trait.df <- data.frame(TRAITDATA,row.names=TRAITDATA$species_ID) 

# Subset the relevant columns
trait.df <- trait.df[,c(4:6)]

#Remove rows that are not present in INSECTDATA
# Remove rows from trait.df where species_ID is not a column in INSECTDATA
species_in_insectdata <- colnames(INSECTDATA)
trait.df <- trait.df[rownames(trait.df) %in% species_in_insectdata, ]

# Convert specific columns to factors
trait.df$diet<-as.factor(trait.df$diet)
trait.df$size<-as.factor(trait.df$size)
trait.df$native<-as.factor(trait.df$native)
```

```{r}
#STEP 0.4.2: Create beetle.df
beetle.df <- data.frame(INSECTDATA,row.names=INSECTDATA$site_ID)
```

######################## 

## ENVIRONMENTAL VARIABLE CORRELATION

######################## 

```{r}
#STEP 0.5.1: Tesing correlation between environmental variables
#Create site.df with row names and select columns 4:32
site.df <- data.frame(ENVIRONMENTDATA, row.names = ENVIRONMENTDATA$site_ID)

# Select only numeric columns from the specified range
site.df <- site.df[, c(4:30)]
site.df <- site.df[sapply(site.df, is.numeric)]

# Now calculate the Pearson correlation matrix on the filtered data
cor_matrix <- cor(site.df, method = "pearson")

# Check for high positive correlations (> 0.6)
high_positive_cor <- cor_matrix > 0.6

# Check for strong negative correlations (< -0.6) with the first column removed
strong_negative_cor <- cor(site.df[-1], method = "pearson") < -0.6

# Write the correlation matrix to a CSV file
write.csv(cor_matrix, file = "correlation_matrix.csv", row.names = TRUE)

# Write the high positive correlations to a CSV file
write.csv(high_positive_cor, file = "high_positive_correlations.csv", row.names = TRUE)

# Write the strong negative correlations to a CSV file
write.csv(strong_negative_cor, file = "strong_negative_correlations.csv", row.names = TRUE)
```

Notes: Removing all the plant occurrences because the standard
deviations are very low.

############################# 

## SPATIAL AUTOCORRELATION

############################# 

```{r}
#STEP 0.5.1: Testing spatial autocorrelation between sites
#creates new dataframe
spp.rich <- beetle.df %>% transmute(site_ID, richness =  rowSums(.[2:ncol(.)] > 0))

# Subset the data frames
rich.df <- spp.rich[1:62,]
coord.df <- ENVIRONMENTDATA[1:62, 2:3]

#Convert coord.df to an sf object
coord.sf <- st_as_sf(coord.df, coords = c("long", "lat"), crs = 4326)

#Calculate the k-nearest neighbors
knn <- knearneigh(st_coordinates(coord.sf), k = 4)

#Convert the knn object to a neighbor list
nb <- knn2nb(knn)

#Create a spatial weights object
lw <- nb2listw(nb, style = "W", zero.policy = TRUE)

#Now lw contains the spatial weights ready for spatial autocorrelation analysis
```

```{r}
# STEP 0.5.2: Calculates spatial autocorrelation, prepares the data for plotting, create a correlogram plot
correlog_plot <- function(mod, title) {
  # Ensure the showtext package is installed and load it
  if (!requireNamespace("showtext", quietly = TRUE)) {
    install.packages("showtext")
  }
  library(showtext)
  
  # Add Open Sans font
  font_add_google("Open Sans", "open_sans")
  showtext_auto()
  
  CN.SAC <- correlog(x = coord.df$lat, y = coord.df$long, z = mod$residuals, 
                     increment = 2, resamp = 500, latlon = TRUE, na.rm = TRUE, quiet = FALSE)
  
  CN.RESIDUAL <- as.data.frame(cbind(CN.SAC$n, CN.SAC$mean.of.class, CN.SAC$correlation, CN.SAC$p))
  colnames(CN.RESIDUAL) <- c("n", "mean", "moransI", "p")
  
  ggplot(CN.RESIDUAL, aes(mean, moransI)) + 
    geom_hline(aes(yintercept = 0), col = "gray30", linetype = "dashed", linewidth = 0.5) +
    geom_line(color = "grey", linewidth = 1) + 
    geom_point(aes(colour = cut(p, c(0, 0.05, 1))), linewidth = 2) +
    scale_color_manual(values = c("(0,0.05]" = "red", "(0.05,1]" = "gray60")) + 
    ylim(-1.5, 1.5) + 
    theme_classic(base_size = 80, base_family = "open_sans") + 
    theme(legend.position = "none") +
    ylab("Moran's I") + 
    xlab("Distance (km)") + 
    ggtitle(title)
}
```

```{r}
# STEP 0.5.3: Spatial autocorrelation analysis
rich.mod <- lm(richness ~ 1, data=rich.df)
lm.morantest(rich.mod, lw, alternative = "two.sided")
tasrich <- correlog_plot(rich.mod, "Tasmania leaf litter beetles - Richness Model")

#Manually perform the Moran's I test to inspect
morans_result <- lm.morantest(rich.mod, lw, alternative = "two.sided")
print(morans_result)
```

```{r}
#STEP 0.5.4: Create grid species plot
tiff("C:\\Users\\Tessas\\OneDrive - University of Tasmania\\Tessa_PhD\\WORK_Code\\CH2_LOCALSCALE\\CH2_4thCornerAnalysis_20241114\\beetlerichness_spatcor_20240822.tiff", units="in", width=18, height=9, res=300, bg = "transparent")
spatplots <- grid.arrange(tasrich)
```

############################# 

## TRAIT-BASED MODEL

############################# 

```{r}
#STEP 0.6.1: Ensure all species in trait.df are present in tas.beetle by padding with zeros
all_species <- rownames(trait.df) #Extracts all species names from the rows of the trait.df dataframe (which contains species traits).
for (species in setdiff(all_species, colnames(beetle.df))) {
  beetle.df[[species]] <- 0 } # for each species that is present in trait.df but missing in tas.beetle, it adds a new column in tas.beetle filled with zeros, ensuring that both datasets have the same species. This step handles cases where some species have traits but are not observed in the tas.beetle dataset (presumably an abundance or occurrence dataset).

# Align species order in tas.beetle to match trait.df
beetle.df <- beetle.df[, all_species] #Rearranges the columns of the tas.beetle dataframe to match the order of species in trait.df. This step ensures that both datasets are aligned properly before running the model.

# Run the trait-environment model
traitmodel <- traitglm(beetle.df, site.df, trait.df, method = "glm1path", family = "negative.binomial") #Fits a Generalized Linear Model (GLM) to analyze the relationship between species abundance (from tas.beetle), environmental variables (from Tas.site), and species traits (from trait.df). The model uses a negative.binomial family to account for overdispersion in count data, and the method glm1path indicates a specific approach within the GLM framework.

# Plot the results
plot(traitmodel)  # Generates a Dunn-Smyth residual plot to check the fit of the model. This plot helps assess the quality of the model by showing the residuals.

# Extract residuals or coefficients
summary(traitmodel) # View model summary for coefficients and p-values

# Normal quantile plot
qqnorm(residuals(traitmodel)) #Creates a normal quantile plot of the model residuals to assess whether the residuals follow a normal distribution, which is an important diagnostic check for model assumptions.
abline(c(0, 1), col = "red") #Adds a red line with slope 1 and intercept 0 to the quantile plot for reference, indicating the expected distribution if the residuals are normally distributed.
```

```{r}
#STEP 0.6.2: Extract the fourth-corner results (assuming it's already a data frame or matrix)
trait.tasbeetles <- traitmodel$fourth.corner #Extracts the fourth-corner results from the traitmodel. The fourth-corner analysis is used to examine the relationships between species traits and environmental variables.

# Check the structure of the results
str(trait.tasbeetles) #Checks the structure of the trait.tasbeetles object, giving an overview of its dimensions, variable types, and content.

# Optional: Save the results as a CSV file
write.csv(trait.tasbeetles, 'C:\\Users\\Tessas\\OneDrive - University of Tasmania\\Tessa_PhD\\WORK_Code\\CH2_LOCALSCALE\\CH2_4thCornerAnalysis_20241114\\trait.response_TasBeetles.csv') #Exports the fourth-corner analysis results to a CSV file for further analysis or reporting.
```

############################# 

## SPECIES-BASED MODEL

############################# 

#### #Run 1, using selection of 22 species

```{r}
# STEP 0.7.1: Check the dataset before starting modelling

# Calculate the correlation matrix for the dataset `site.df`
# `use = "complete.obs"` ensures that pairwise correlations ignore missing values.
cor_matrix2 <- cor(site.df, use = "complete.obs")

# Save the calculated correlation matrix to a CSV file for further inspection or documentation.
write.csv(cor_matrix2, "correlation_matrix2.csv", row.names = TRUE)
```

```{r}
# STEP 0.7.2: Clean the data
# Define a custom function to remove rows with NA values from a dataframe
remove_na_rows <- function(df, df_name) {
  # Identify columns containing missing (NA) values
  na_columns <- colnames(df)[colSums(is.na(df)) > 0]
  
  # Notify the user about columns with missing values
  if (length(na_columns) > 0) {
    cat(paste0("Columns with missing values in '", df_name, "' removed:\n"))
    print(na_columns)
  } else {
    # Notify if there are no columns with missing values
    cat(paste0("No columns with missing values found in '", df_name, "'.\n"))
  }
  
  # Remove rows containing any NA values
  df_clean <- na.omit(df)
  
  # Inform the user about the number of rows retained after cleaning
  cat(paste0(nrow(df_clean), " rows retained in '", df_name, "' after removing NAs.\n"))
  
  return(df_clean)
}

# Apply the function to both `site.df` and `beetle.df` datasets
site.df <- remove_na_rows(site.df, "site.df")
beetle.df <- remove_na_rows(beetle.df, "beetle.df")

# Remove non-numeric columns from `beetle.df`
beetle.df <- beetle.df[sapply(beetle.df, is.numeric)]

# Remove non-numeric columns from `site.df`
site.df <- site.df[sapply(site.df, is.numeric)]
```

```{r}
# STEP 0.7.3: Filter the beetle dataset to include only specified species
# List of species to retain
selected_species <- c(
  "Sloaneana_tasmaniae", "Geomela_bryophaga", "Pseudostanus_tasmanicus", "Cartodere_australica",
  "Cryptamorpha_triguttata", "Xynotropis_TFIC_sp_01", "Mandalotus_muscivorus", "Decilaus_nigronotatus",
  "Decilaus_striatus", "Decilaus_bryophilus", "Exithius_TFIC_sp_01", "Lophixus_major",
  "Notoceryon_TFIC_sp_B", "Saprosites_mendax", "Nargomorphus_globulus", "Ptinella_TFIC_SP_01",
  "Quedius_inaqualipennis", "Tasmanityrus_newtoni", "Anepius_koebelei", "Orchesia_pictipennis",
  "Diemenoma_commoda", "Enhypnon_latitarsis"
)

# Filter the beetle dataset
beetle.df <- beetle.df[, colnames(beetle.df) %in% selected_species]

# Ensure `beetle.df` contains integer values
beetle.df <- round(beetle.df)

# Inform the user about rounding
cat("Rounded `beetle.df` to integer values for compatibility with count-based models.\n")

# Inform the user about the filtering step
cat("Filtered `beetle.df` to include only the specified species.\n")
```

```{r}
# STEP 0.7.4: Trait-based modelling
# Attempt the Negative Binomial model
sppmodel <- tryCatch({
  traitglm(beetle.df, site.df, method = "glm1path", family = "negative.binomial")
}, error = function(e) {
  cat("Error with Negative Binomial model: ", e$message, "\n")
  NULL
})

# Check if model fit successfully and plot residuals
if (!is.null(sppmodel)) {
  plot(sppmodel) # Dunn-Smyth residual plot
} else {
  # Try the Poisson model if the Negative Binomial model fails
  sppmodel <- tryCatch({
    traitglm(beetle.df, site.df, method = "glm1path", family = "poisson")
  }, error = function(e) {
    cat("Error with Poisson model: ", e$message, "\n")
    NULL
  })
  
  if (!is.null(sppmodel)) {
    plot(sppmodel)
  }
}

# Check normality of residuals if model fits
if (!is.null(sppmodel)) {
  qqnorm(residuals(sppmodel)); abline(c(0, 1), col = "red")

# Ensure output directory exists
output_dir <- "C:/Users/Tessas/OneDrive - University of Tasmania/Tessa_PhD/WORK_Code/CH2_LOCALSCALE/CH2_4thCornerAnalysis_20241114"
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
  cat("Created output directory:", output_dir, "\n")
} else {
  cat("Output directory already exists:", output_dir, "\n")
}

# Export species response data
output_file <- file.path(output_dir, "species.response_TasBeetles.csv")
cat("Attempting to write to file:", output_file, "\n")

tryCatch({
  write.csv(spp.tasbeetles, output_file)
  cat("File successfully written to:", output_file, "\n")
}, error = function(e) {
  cat("Error writing file:", e$message, "\n")
})
```

############################## 

## BIOMASS ESTIMATION

```{r}
# STEP 0.8.1: Assuming TRAITDATA is already loaded
# Check if 'size.1' exists in the dataframe before proceeding
if ("size.1" %in% colnames(TRAITDATA)) {
  # Add a new column 'Biomass_modifier' based on the condition of 'size.1'
  TRAITDATA_biomass <- TRAITDATA %>%
    mutate(Biomass_modifier = ifelse(size.1 < 10, 0.0020, 0.024))
  
  # Display the updated dataframe
  head(TRAITDATA_biomass)
} else {
  # Print a message if 'size.1' does not exist
  cat("Column 'size.1' does not exist in the TRAITDATA dataframe.\n")
}
```

```{r}
# STEP 0.8.2: Convert INSECTDATA to long format
# Assuming `INSECTDATA` has site_ID as row identifiers and species as column headers
insectdata_long <- INSECTDATA %>%
  pivot_longer(
    cols = -site_ID,                # Exclude site_ID from pivoting
    names_to = "species_ID",        # Create a column for species IDs
    values_to = "count"             # Create a column for counts
  )

# Display the transformed data
head(insectdata_long)
```

```{r}
# STEP 0.8.3: Merge insectdata_long with TRAITDATA_biomass using species_ID
merged_data <- insectdata_long %>%
  left_join(TRAITDATA_biomass, by = "species_ID")

# Define a function to calculate biomass based on family and size
calculate_biomass <- function(family, length, count) {
  # Handle cases where length is missing
  if (is.na(length)) {
    return(NA)
  }
  
  if (is.na(family)) {
    # Handle missing family values
    if (length < 10) {
      biomass <- 0.0020 * (length ^ 3.17) * count
    } else {
      biomass <- 0.024 * (length ^ 2.92) * count
    }
  } else if (family == "Carabidae") {
    biomass <- 0.028 * (length ^ 2.75) * count
  } else if (family == "Staphylinidae") {
    biomass <- 0.015 * (length ^ 3.10) * count
  } else if (length < 10) {
    biomass <- 0.0020 * (length ^ 3.17) * count
  } else {
    biomass <- 0.024 * (length ^ 2.92) * count
  }
  
  return(biomass)
}

# Apply the biomass calculation for each row
merged_data <- merged_data %>%
  rowwise() %>%
  mutate(
    Biomass = calculate_biomass(
      family = family,  # Use 'family' for the family column
      length = size.1,  # Use 'size.1' for beetle length
      count = count     # Use 'count' for the species count
    )
  ) %>%
  ungroup()

# Summarise total biomass by site_ID
biomass_by_site <- merged_data %>%
  group_by(site_ID) %>%
  summarise(
    total_biomass = sum(Biomass, na.rm = TRUE)
  )

# Save summarised biomass by site_ID to a CSV file
write.csv(biomass_by_site, "total_biomass_by_site.csv", row.names = FALSE)

# Display a message to confirm the file was saved
cat("Summed biomass by site saved to 'total_biomass_by_site.csv'\n")
```

```{r}
# STEP 0.8.4: Load the biomass and abundance dataset
Biomass_area <- read.csv("biomass_by_site_abundance_area.csv", header = TRUE, stringsAsFactors = FALSE)

# Check the structure of the dataset
print(head(Biomass_area))
print(summary(Biomass_area))

# Ensure 'type' and 'amount' columns exist
if(!"type" %in% colnames(Biomass_area) | !"amount" %in% colnames(Biomass_area)) {
  stop("The columns 'type' or 'amount' are missing from the Biomass_area dataset.")
}

# Convert 'amount' to numeric, handling any non-numeric values
Biomass_area$amount <- as.numeric(Biomass_area$amount)

# Check for any NA values introduced during conversion
print(sum(is.na(Biomass_area$amount)))

# Reshape the data: Pivot 'type' column to create 'total_biomass' and 'beetles_m2' columns
library(tidyr)
Biomass_area_wide <- Biomass_area %>%
  pivot_wider(
    names_from = type,
    values_from = amount,
    values_fill = list(amount = NA)  # Fill missing values with NA
  )

# Verify the reshaped data
print(head(Biomass_area_wide))

# Optional: View the reshaped dataset
view(Biomass_area_wide)
```

```{r}
# STEP 0.8.5: Merge the forest_type column from ENVIRONMENTDATA to Biomass_area_wide using site_ID as the key
Biomass_area_wide <- merge(
  x = Biomass_area_wide,
  y = ENVIRONMENTDATA,
  by = 'site_ID',
  all.x = TRUE
)

# Ensure total_biomass and beetles_per_area are numeric, and forest_type is categorical
Biomass_area_wide$total_biomass <- as.numeric(as.character(Biomass_area_wide$total_biomass))
Biomass_area_wide$beetles_per_area <- as.numeric(as.character(Biomass_area_wide$beetles_per_area))
Biomass_area_wide$forest_type <- factor(Biomass_area_wide$forest_type, levels = c("rainforest", "wet_mixed", "wet"))

# Keep rows with values but ignore missing data for each measurement during plotting
library(tidyr)
Biomass_area_long <- Biomass_area_wide %>% 
  pivot_longer(cols = c(total_biomass, beetles_per_area), names_to = "measurement", values_to = "value")

# Keep rows with non-missing values for either total_biomass or beetles_per_area and only specific forest types
Biomass_area_long_plot <- Biomass_area_long %>% 
  filter(!is.na(value) & forest_type %in% c("rainforest", "wet_mixed", "wet"))

# Display the result
print(Biomass_area_wide)
```

```{r}
# STEP 0.8.6: Plot boxplots for Biomass and Beetles per Area by Forest Type, excluding rows with NA in the plotted measurement or forest_type
ggplot(Biomass_area_long_plot, aes(x = forest_type, y = value, fill = forest_type)) +
  geom_boxplot(color = "black") +
  facet_wrap(~measurement, scales = "free_y", labeller = as_labeller(c(
    total_biomass = "Biomass (mg/m²)",
    beetles_per_area = "Abundance/m²"
  ))) +
  scale_y_continuous(limits = c(0, 450)) +
  scale_fill_manual(values = c("#3366FF", "#009999", "#009999")) + # Added color for "Mixed"
  theme_minimal() +
  theme(
    strip.text = element_text(face = "bold", size = 12),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank(), # Remove grid lines
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1) # Add boxes around facets
  ) +
  xlab("") + # Remove x-axis label
  ylab("") # Remove the y-axis label
```

END DOCUMENT
