---
title: "CH3_Nestedness"
author: "Tessa R Smith"
date: "2024-03-01"
output: html_document
---

## INTRODUCTION

**BEETLE BIOGEOGRAPHY PROJECT**, University of Tasmania. Data collected by Tessa R Smith and volunteers (2020-2022). Code by Tessa Smith, last edited December 2024.

**AIMS**
-   **AIM 2:Exploratory data analysis, Nestedness analysis by site**

## DATA PREPARATION

```{r, echo = FALSE}
#STEP 0.1: Setup packages and working directory
rm(list = ls()) # clear memory and start afresh

#Change the location for your computer. 
#NOTE make sure you use \\ instead of \ or /, as these result in errors. 
setwd("C:\\Users\\Tessas\\OneDrive - University of Tasmania\\Tessa_PhD\\WORK_Code\\CH3_BROADSCALE\\CH3_analysis5_Nestedness")
```
Notes: works

```{r}
#Load packages
library(dplyr)
library(tidyverse)
```
Notes: works

```{r, echo = FALSE}
#STEP 0.2: Import data
#Metadata (includes altitude etc)
#Reference: Own data (Tessa Smith PhD)
#File type: CSV
#Format: long form. #Columns: site_ID, species, specimen_count. #Rows: Individual records
beetle_records.df <- read.csv("TS_Site_BeetleRecords_20240306.csv", header = TRUE, stringsAsFactors = FALSE)
View(beetle_records.df) #Opens data in another tab so it can be easily viewed
```
Notes: works

```{r, echo = FALSE}
#STEP 0.2.2: Reshapes species list of beetle records from long form into site.count form (wide)
#Only necessary once before start of analysis
df_reshaped <- beetle_records.df %>%
#Change from long to wide format
pivot_wider(id_cols = site_ID,
            names_from = species,
            values_from = species_count,
            # Sum the entries where there is more than 1 value in specimen_count
            values_fn = list(species_count= sum)) 

#STEP 0.2.4: Change NAs in the new dataset to 0
df_reshaped[is.na(df_reshaped)] <-0 
df_reshaped <- na.omit(df_reshaped)

#STEP 0.2.5: Add new columns, Count abundance number for each site, Cound species number for each site, Count unique species for each site
df_reshaped <- df_reshaped %>% 
  mutate(
    abundance = rowSums(.[, -1]),  # Sum all values from columns other than the first one
    species_count = rowSums(.[-1] > 0),  # Count numbers of columns >0
    unique_species = rowSums((.[, -1] > 0) * (colSums(.[-1] > 0) == 1))  # Count number of columns with a value >0 that occur in no other row
  )
view (df_reshaped)
```
Notes: works

```{r, echo = FALSE}
#STEP 0.3: Import data (environmental data(local))
#Metadata (includes altitude etc)
#Reference: Own data (Tessa Smith PhD)
#File type: CSV
#Format: wide form #Columns: environmental variables #Rows:Sites
Metadata <- read.csv("TS_Site_Enviro_metadata_20230622.csv", header = TRUE, stringsAsFactors = FALSE)
View(Metadata) #Opens data in another tab so it can be easily viewed

```
Notes: works

Notes:
IBRA Bioregion colours
TSR01-#fde725
TSE01-#a0da39
BEL01- #4ac16d
FUR01- #1fa187
TNS01-#277f8e
KIN01- #365c8d
TWE01-#46327e
TCH01-#440154

**AIM 5:Nestedness analysis by site**
Code source: https://nilanjanchatterjee.github.io/com_eco/
Note: Code not updated from original code

Calculation of nestedness, the value ranges from 0-100 where 0 stands for perfectly nested community and 100 stands for random community. 
```{r}
#STEP 5.1.1: Load packages
library(vegan)
```

```{r}
#STEP 5.2: Calculate nestedness
nesttry<-nestedtemp(df_reshaped[, -1])
nesttry
```

```{r}
#STEP 5.3: Plot nestedness
plot(nesttry, xlab = "Species", ylab="Sites",main="Extinction probability")
plot(nesttry, kind="incidence",xlab = "Species", ylab="Sites",main="Presence-absence")
```
