---
title: "CH3_NMDS"
author: "Tessa R Smith"
date: "2024-03-01"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r, echo = FALSE}
#STEP 0.1.1: Setup packages and working directory
rm(list = ls()) # clear memory and start afresh

#Change the location for your computer. 
#NOTE make sure you use \\ instead of \ or /, as these result in errors. 
setwd("C:\\Users\\Tessas\\OneDrive - University of Tasmania\\Tessa_PhD\\WORK_Code\\CH3_BROADSCALE\\CH3_NMDS_20241114")
```

# INTRODUCTION

**BEETLE BIOGEOGRAPHY PROJECT**, University of Tasmania. Data collected
by Tessa R Smith and volunteers (2020-2022). Code by Tessa Smith, last
edited December 2024.

-   **AIMS**

-   **AIM 1:Exploratory data analysis, NMDS of beetle records by site**

-   **AIM 2: Descriptive statistics, Diversity with altitude**

## DATA PREPARATION

```{r}
#STEP 0.1.2: Load packages
library(dplyr) # For data manipulation
library(tidyverse)
library(vegan)
library(ggplot2)
library(svglite)
library(ggpubr)  # For stat_regline_equation()
library(sf)
```

Data pre treatment: 1. Creating a species column with no spaces and
without underscore at end of text -create new column for species -TRIM
(removes all spaces except one between words), data from species column
-copy and paste values only -replace all the spaces in the new species
column with underscore (\_) -delete original species column -rename new
species column

2.  Delete any rows without data -conditional formatting for rows with
    blanks -delete these rows

3.  Remove species types that you don't want "UNKNOWN_SPECIES",
    "UNKNOWN_SPECIES-Pieces", "PELORIDIIDAE_indet.", "Hemiodoecus_leai",
    "Hemiodoecellus_fidelis", "Choerocoris_paganus",
    "LYCIDAE_indet.\_(larvae)", "Misc.\_beetles", "Hemiptera\_(misc.)",
    "Spiders\_(misc.)", "Opiliones\_(misc.)", "Mites\_(misc.)",
    "Collembola\_(misc.)", "Millipedes\_&*entipedes (misc.)",
    "Earwigs*(misc.)", "Hymenoptera\_(misc.)", "Snails\_(misc.)",
    "Other_non-beetle_invertebrates\_(misc.)"

```{r, echo = FALSE}
#STEP 0.2.1: Import data
beetle_records.df <- read.csv("TS_Site_BeetleRecords_20241212.csv", header = TRUE, stringsAsFactors = FALSE)
View(beetle_records.df) #Opens data in another tab so it can be easily viewed
```

Notes: works

```{r, echo = FALSE}
# STEP 0.2.2: Reshape species list of beetle records from long form into site.count form (wide)
# Only necessary once before start of analysis
df_reshaped <- beetle_records.df %>%
  # Remove rows with no entry in the species column
  filter(!is.na(species) & species != "") %>%
  # Change from long to wide format
  pivot_wider(
    id_cols = site_ID,
    names_from = species,
    values_from = species_count,
    # Sum the entries where there is more than 1 value in specimen_count
    values_fn = list(species_count = sum)
  )

# Change NAs in the new dataset to 0
df_reshaped[is.na(df_reshaped)] <- 0 

# Add new columns
# Count abundance number for each site, count species number for each site, count unique species for each site
df_reshaped <- df_reshaped %>% 
  mutate(
    abundance = rowSums(.[, -1]),  # Sum all values from columns other than the first one
    Count = rowSums(.[-1] > 0),  # Count numbers of columns >0
    unique_species = rowSums((.[, -1] > 0) * (colSums(.[-1] > 0) == 1))  # Count number of columns with a value >0 that occur in no other row
  )

# View the resulting dataframe
view(df_reshaped)
```

```{r}
#STEP 0.2.3: Remove columns that we don't want in the database

#create list of names of columns that we want to remove

columns_to_remove <- c("UNKNOWN_SPECIES",
"UNKNOWN_SPECIES-Pieces",
"PELORIDIIDAE_indet.",
"Hemiodoecus_leai",
"Hemiodoecellus_fidelis",
"Choerocoris_paganus",
"LYCIDAE_indet._(larvae)",
"Misc._beetles",
"Hemiptera_(misc.)",
"Spiders_(misc.)",
"Opiliones_(misc.)",
"Mites_(misc.)",
"Collembola_(misc.)",
"Millipedes_&_entipedes (misc.)",
"Earwigs_(misc.)",
"Hymenoptera_(misc.)",
"Snails_(misc.)",
"Other_non-beetle_invertebrates_(misc.)", 
"STAPHYLINIDAE_indet.",
"CURCULIONIDAE_indet.")

# remove the columns
# Remove the columns from the dataframe
df_reshaped <- df_reshaped[ , !(names(df_reshaped) %in% columns_to_remove)]

# View the updated dataframe
View(df_reshaped)
```

```{r}
#STEP 0.2.4: Log transform data
for (i in 2:675) {
  df_reshaped[, i] = log(df_reshaped[, i] + 1)
}
head(df_reshaped)
```

Notes: works. Logarithmic transformation (log(x + 1)) is frequently used
to stabilize variance and normalize the distribution of species
abundance data. It is particularly useful when dealing with highly
skewed data, as it compresses the range of values and reduces the impact
of large outliers.

```{r}
#STEP 0.2.5: Remove rare species (species present in <5 sites)

# Function to remove columns where species is present in less than 5 sites
remove_infrequent_species <- function(df, threshold = 5) {
  # Replace any " in column names with _
  colnames(df) <- gsub('"', '_', colnames(df))
  
  # Calculate the number of sites each species is present in
  species_counts <- colSums(df != 0)
  
  # Identify columns to keep and to remove
  columns_to_keep <- species_counts >= threshold
  columns_to_remove <- species_counts < threshold
  
  # Filter the data frame
  df_filtered <- df[, columns_to_keep]
  
  # Print the number of columns included and removed
  num_included <- sum(columns_to_keep)
  num_removed <- sum(columns_to_remove)
  cat("Number of columns included:", num_included, "\n")
  cat("Number of columns removed:", num_removed, "\n")
  
  return(df_filtered)
}

# Apply the function to df_reshaped
df_filtered <- remove_infrequent_species(df_reshaped, 5)
```

Notes: works. Removes the majority of the data though.

\*Skipping this step for run on 26/08/2024

```{r}
#STEP 0.2.5.1: Run if skipping previous step
df_filtered <-df_reshaped
```

```{r, echo = FALSE}
#STEP 0.3.1: Import data (environmental data(local))
#Metadata (includes altitude etc)
#Reference: Own data (Tessa Smith PhD)
#File type: CSV
#Format: wide form #Columns: environmental variables #Rows:Sites
Metadata <- read.csv("TS_Site_Enviro_metadata_20230622.csv", header = TRUE, stringsAsFactors = FALSE)
View(Metadata) #Opens data in another tab so it can be easily viewed
```

## **AIM 1:Exploratory data analysis, NMDS of beetle records by site** 

Method reference: MultiDimensional Scaling (NMDS) is a distance-based
ordination technique.
<https://uw.pressbooks.pub/appliedmultivariatestatistics/chapter/nmds/>


```{r, echo = FALSE}
#STEP 1.1.1: Add IBRA Bioregion to Records_dataset
# Perform a left join to align the dataframes
Records_data <- left_join(df_filtered, Metadata, by = "site_ID")

#Convert site_ID names to numeric values
Records_data$site_ID <- as.numeric(factor(df_filtered$site_ID))
head(Records_data) #Reminder to convert this back again later!
```

```{r}
#STEP 1.1.2: Select columns 1 to 735 from the 'Records_data' dataframe
Records_data_species <- Records_data %>% select(1:134)

# View the resulting dataframe
View(Records_data_species)
```

```{r}
#STEP 1.1.3: Run the NMDS
set.seed(75) 

z_beetles <- metaMDS (comm = Records_data_species,
autotransform = FALSE, #If a transformation is applied to the data #CHANGED FROM TRUE
distance = "bray", #Bray-curtis dissimilarity metric
engine = "monoMDS", #the default engine in vegan
k = 3, #number of dimensions in the solution                        
weakties = TRUE, #weak ties should be allowed in the rank order of dissimilarities
model = "global",  #Global model (This is the normal non-metric MDS with a monotone regression. The global model attempts to find a configuration of points in low-dimensional space (usually 2D or 3D) that preserves the rank order of distances between all pairs of objects in the original high-dimensional space) vs. local model (This is a non-metric MDS with separate regressions for each point. The local model allows for different monotonic relationships between observed and predicted dissimilarities for different objects. This can be useful when the relationship between objects varies across the gradient). Other options are linear and hybrid models. 
maxit = 400, #Maximum number of permitted iterations
try = 40,  #Number of random starts in the search for a stable solution
trymax = 100)

# Plot goodness of fit
gof <- goodness(object = z_beetles)
plot(z_beetles, display = "sites", type = "none")

points(z_beetles, display = "sites",
cex = 2*gof/mean(gof))

#Shephard plot
plot(z_beetles$diss, z_beetles$dist)

# Calculate species scores as the weighted average of the coordinates of the plots on which those species occurred.

# Obtain ordination coordinates
z_beetles$points %>% head()

z.points <- data.frame(z_beetles$points)
```

Notes: x is observed dissimilarity y is ordination distance Notes: works

```{r}
#STEP 1.1.4: NMDS run without ALA data

# Remove `site_ID` or any non-numeric columns
df_filtered_numeric <- df_filtered %>%
  select(-site_ID)  # Ensure site_ID is excluded

# Convert to a numeric matrix
comm_matrix <- as.matrix(df_filtered_numeric)

# Replace missing values (NA) with 0
comm_matrix[is.na(comm_matrix)] <- 0

# Ensure all entries are numeric
if (!all(sapply(comm_matrix, is.numeric))) {
  stop("Error: Non-numeric values detected in the matrix. Please check your data.")
}

# Run the NMDS (no ala data)
set.seed(12345) 
z_beetles <- metaMDS (comm = df_filtered_numeric,
autotransform = FALSE, #If a transformation is applied to the data #already transformed
distance = "bray", #Bray-curtis dissimilarity metric
engine = "monoMDS", #the default engine in vegan
k = 2, #number of dimensions in the solution                        
weakties = TRUE, #weak ties should be allowed in the rank order of dissimilarities
model = "global",  #Global model (This is the normal non-metric MDS with a monotone regression. The global model attempts to find a configuration of points in low-dimensional space (usually 2D or 3D) that preserves the rank order of distances between all pairs of objects in the original high-dimensional space) vs. local model (This is a non-metric MDS with separate regressions for each point. The local model allows for different monotonic relationships between observed and predicted dissimilarities for different objects. This can be useful when the relationship between objects varies across the gradient). Other options are linear and hybrid models. 
maxit = 200, #Maximum number of permitted iterations
try = 20,  #Number of random starts in the search for a stable solution
trymax = 100)

# Plot goodness of fit
gof <- goodness(object = z_beetles)
plot(z_beetles, display = "sites", type = "none")

points(z_beetles, display = "sites",
cex = 2*gof/mean(gof))

#Shephard plot
plot(z_beetles$diss, z_beetles$dist)

# Calculate species scores as the weighted average of the coordinates of the plots on which those species occurred.

# Obtain ordination coordinates
z_beetles$points %>% head()

z.points <- data.frame(z_beetles$points)

```

```{r}
#STEP 1.1.5: Merge Records_data with z.points
z.points <- cbind(z.points, Records_data)
```


```{r}
#STEP 1.2.1: Plot NMDS by IBRA Bioregion

# Merge z.points with Records_data to include IBRA bioregion and forest type
z.points <- z.points %>%
    mutate(ibra_region = Records_data$ibra_region, 
           forest_type = Records_data$forest_type)

# Preprocess z.points to remove NAs and group by IBRA bioregion
z.points_clean <- z.points %>%
    filter(!is.na(ibra_region))  # Exclude NAs in IBRA bioregion

# Calculate convex hulls for each IBRA bioregion
convex_hulls <- z.points_clean %>%
    group_by(ibra_region) %>%
    slice(chull(MDS1, MDS2)) %>%
    ungroup()

# Convert convex hulls back to a data frame for ggplot
convex_hulls <- as.data.frame(convex_hulls)

# Plot NMDS by IBRA Bioregion with convex hulls
NMDS_IBRA <- ggplot(z.points_clean, aes(x = MDS1, y = MDS2)) + 
    # Add convex hull polygons
    geom_polygon(data = convex_hulls, aes(x = MDS1, y = MDS2, fill = ibra_region, group = ibra_region), 
                 alpha = 0.3, colour = "black") +
    # Add points
    geom_point(size = 3, aes(shape = forest_type, colour = ibra_region)) + 
    # Custom theme
    theme(axis.text.y = element_text(colour = "black", size = 10, face = "bold"), 
          axis.text.x = element_text(face = "bold", size = 10, colour = "black"), 
          legend.text = element_text(size = 10, face = "bold", colour = "black"), 
          legend.position = "right", 
          axis.title.y = element_text(face = "bold", size = 10), 
          axis.title.x = element_text(face = "bold", size = 10, colour = "black"), 
          legend.title = element_text(size = 10, colour = "black", face = "bold"), 
          panel.background = element_blank(), 
          panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
          legend.key = element_blank()) + 
    # Labels and colours
    labs(x = "NMDS1", colour = "IBRA Bioregion", y = "NMDS2", shape = "Forest Type", fill = "IBRA Bioregion") +
    scale_colour_manual(values = c("#4ac16d", "#1fa187", "#365c8d", "#440154", "#277f8e", "#a0da39", "#fde725", "#46327e", "#E69F00")) +
    scale_fill_manual(values = c("#4ac16d", "#1fa187", "#365c8d", "#440154", "#277f8e", "#a0da39", "#fde725", "#46327e", "#E69F00"))  # Match fill colours to point colours

# Display the plot
NMDS_IBRA

# Save the plot as a .png file
ggsave("NMDS_IBRA.png", plot = NMDS_IBRA, width = 8, height = 6, dpi = 300)

```

```{r}
#STEP 1.2.2: Plot NMDS by Ecoregions
# Merge z.points with Records_data to include ecoregion and forest_type
z.points <- z.points %>%
    mutate(ecoregion = Records_data$ecoregion, 
           forest_type = Records_data$forest_type)

# Preprocess z.points to remove NAs and group by ecoregion
z.points_clean <- z.points %>%
    filter(!is.na(ecoregion))  # Exclude NAs in ecoregion

# Calculate convex hulls for each ecoregion
convex_hulls <- z.points_clean %>%
    group_by(ecoregion) %>%
    slice(chull(MDS1, MDS2))  # Compute convex hull points for each ecoregion

# Plot NMDS by Ecoregion with convex hulls
NMDS_Ecoregion <- ggplot(z.points_clean, aes(x = MDS1, y = MDS2)) + 
    # Add convex hull polygons
    geom_polygon(data = convex_hulls, aes(x = MDS1, y = MDS2, fill = ecoregion, group = ecoregion), 
                 alpha = 0.3, colour = "black") +
    # Add points
    geom_point(size = 3, aes(shape = forest_type, colour = ecoregion)) + 
    # Custom theme
    theme(axis.text.y = element_text(colour = "black", size = 10, face = "bold"), 
          axis.text.x = element_text(face = "bold", size = 10, colour = "black"), 
          legend.text = element_text(size = 10, face = "bold", colour = "black"), 
          legend.position = "right", 
          axis.title.y = element_text(face = "bold", size = 10), 
          axis.title.x = element_text(face = "bold", size = 10, colour = "black"), 
          legend.title = element_text(size = 10, colour = "black", face = "bold"), 
          panel.background = element_blank(), 
          panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
          legend.key = element_blank()) + 
    # Labels and colours
    labs(x = "NMDS1", colour = "Ecoregion", y = "NMDS2", shape = "Forest Type", fill = "Ecoregion") +
    scale_colour_manual(values = c("#50BF7F", "#31CFB6", "#60B223")) +
    scale_fill_manual(values = c("#50BF7F", "#31CFB6", "#60B223"))  # Match fill colours to point colours

# Display the plot
NMDS_Ecoregion

# Save the plot as a .png file
ggsave("NMDS_Ecoregion.png", plot = NMDS_Ecoregion, width = 8, height = 6, dpi = 300)
```

```{r}
#STEP 1.2.3:  Plot NMDS by Invertebrate Bioregion

# Merge z.points with Records_data to include Invertebrate Bioregion and forest type
z.points <- z.points %>%
    mutate(invert_bioreg = Records_data$invert_bioreg, 
           forest_type = Records_data$forest_type)

# Preprocess z.points to remove NAs and group by Invertebrate Bioregion
z.points_clean <- z.points %>%
    filter(!is.na(invert_bioreg))  # Exclude NAs in Invertebrate Bioregion

# Calculate convex hulls for each Invertebrate Bioregion
convex_hulls <- z.points_clean %>%
    group_by(invert_bioreg) %>%
    slice(chull(MDS1, MDS2)) %>%
    ungroup()

# Plot NMDS by Invertebrate Bioregion with convex hulls
NMDS_InvertBioreg <- ggplot(z.points_clean, aes(x = MDS1, y = MDS2)) + 
    # Add convex hull polygons
    geom_polygon(data = convex_hulls, aes(x = MDS1, y = MDS2, fill = invert_bioreg, group = invert_bioreg), 
                 alpha = 0.3, colour = "black") +
    # Add points
    geom_point(size = 3, aes(shape = forest_type, colour = invert_bioreg)) + 
    # Custom theme
    theme(axis.text.y = element_text(colour = "black", size = 10, face = "bold"), 
          axis.text.x = element_text(face = "bold", size = 10, colour = "black"), 
          legend.text = element_text(size = 10, face = "bold", colour = "black"), 
          legend.position = "right", 
          axis.title.y = element_text(face = "bold", size = 10), 
          axis.title.x = element_text(face = "bold", size = 10, colour = "black"), 
          legend.title = element_text(size = 10, colour = "black", face = "bold"), 
          panel.background = element_blank(), 
          panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
          legend.key = element_blank()) + 
    # Labels and colours
    labs(x = "NMDS1", colour = "Invertebrate Bioregion", y = "NMDS2", shape = "Forest Type", fill = "Invertebrate Bioregion") +
    scale_colour_manual(values = c("#b01c38", "#d62618", "#d62618", "#f0e800", "#f0e800", "#f6f195", 
                                   "#4c10ce", "#042d9b", "#042d9b", "#9b83eb", "#a122a2", "#b62ccb", "#333d37")) +
    scale_fill_manual(values = c("#b01c38", "#d62618", "#d62618", "#f0e800", "#f0e800", "#f6f195", 
                                 "#4c10ce", "#042d9b", "#042d9b", "#9b83eb", "#a122a2", "#b62ccb", "#333d37"))  # Match fill colours to point colours

# Display the plot
NMDS_InvertBioreg

# Save the plot as a .png file
ggsave("NMDS_InvertBioreg.png", plot = NMDS_InvertBioreg, width = 8, height = 6, dpi = 300)
```

```{r}
#STEP 1.2.4: Plot NMDS by Tyler's Line

# Merge z.points with Records_data to include Tyler's Line and forest type
z.points <- z.points %>%
    mutate(tylers = Records_data$tylers, 
           forest_type = Records_data$forest_type)

# Preprocess z.points to remove NAs and group by Tyler's Line
z.points_clean <- z.points %>%
    filter(!is.na(tylers))  # Exclude NAs in Tyler's Line

# Calculate convex hulls for each category of Tyler's Line
convex_hulls <- z.points_clean %>%
    group_by(tylers) %>%
    slice(chull(MDS1, MDS2)) %>%
    ungroup()

# Plot NMDS by Tyler's Line with convex hulls
NMDS_Tylers <- ggplot(z.points_clean, aes(x = MDS1, y = MDS2)) + 
    # Add convex hull polygons
    geom_polygon(data = convex_hulls, aes(x = MDS1, y = MDS2, fill = tylers, group = tylers), 
                 alpha = 0.3, colour = "black") +
    # Add points
    geom_point(size = 3, aes(shape = forest_type, colour = tylers)) + 
    # Custom theme
    theme(axis.text.y = element_text(colour = "black", size = 10, face = "bold"), 
          axis.text.x = element_text(face = "bold", size = 10, colour = "black"), 
          legend.text = element_text(size = 10, face = "bold", colour = "black"), 
          legend.position = "right", 
          axis.title.y = element_text(face = "bold", size = 10), 
          axis.title.x = element_text(face = "bold", size = 10, colour = "black"), 
          legend.title = element_text(size = 10, colour = "black", face = "bold"), 
          panel.background = element_blank(), 
          panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
          legend.key = element_blank()) + 
    # Labels and colours
    labs(x = "NMDS1", colour = "Tyler's Line", y = "NMDS2", shape = "Forest Type", fill = "Tyler's Line") +
    scale_colour_manual(values = c("#a83232", "#fcdb03", "#32a855")) +
    scale_fill_manual(values = c("#a83232", "#fcdb03", "#32a855"))  # Match fill colours to point colours

# Display the plot
NMDS_Tylers

# Save the plot as a .png file
ggsave("NMDS_Tylers.png", plot = NMDS_Tylers, width = 8, height = 6, dpi = 300)
```

```{r}
#STEP 1.2.5: Plot NMDS by Geology (age)
# Convert landform_age to a factor with specified levels and labels
Records_data$landform_age <- as.numeric(Records_data$landform_age)
Records_data$landform_age <- factor(Records_data$landform_age, 
                                    levels = c(1, 2, 3, 4), 
                                    labels = c("Cenozoic", "Mesozoic", "Paleozoic", "Precambrian"))

# Align landform_age with z.points
z.points <- z.points %>%
    mutate(landform_age = Records_data$landform_age, 
           forest_type = Records_data$forest_type)

# Preprocess z.points to remove NAs and group by landform_age
z.points_clean <- z.points %>%
    filter(!is.na(landform_age))  # Exclude NAs in landform_age

# Calculate convex hulls for each landform_age
convex_hulls <- z.points_clean %>%
    group_by(landform_age) %>%
    slice(chull(MDS1, MDS2)) %>%
    ungroup()

# Plot NMDS by Geology (landform_age) with convex hulls
NMDS_Geology <- ggplot(z.points_clean, aes(x = MDS1, y = MDS2)) + 
    # Add convex hull polygons
    geom_polygon(data = convex_hulls, aes(x = MDS1, y = MDS2, fill = landform_age, group = landform_age), 
                 alpha = 0.3, colour = "black") +
    # Add points
    geom_point(size = 3, aes(shape = forest_type, colour = landform_age)) + 
    # Custom theme
    theme(
        axis.text.y = element_text(colour = "black", size = 10, face = "bold"), 
        axis.text.x = element_text(face = "bold", size = 10, colour = "black"), 
        legend.text = element_text(size = 10, face = "bold", colour = "black"), 
        legend.position = "right", 
        axis.title.y = element_text(face = "bold", size = 10), 
        axis.title.x = element_text(face = "bold", size = 10, colour = "black"), 
        legend.title = element_text(size = 10, colour = "black", face = "bold"), 
        panel.background = element_blank(), 
        panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
        legend.key = element_blank()
    ) + 
    # Labels and colours
    labs(x = "NMDS1", colour = "Landform Age", y = "NMDS2", shape = "Forest Type", fill = "Landform Age") + 
    scale_colour_manual(values = c("#F5CB5C", "#087CA7", "#6EEB83", "#EA3546")) +
    scale_fill_manual(values = c("#F5CB5C", "#087CA7", "#6EEB83", "#EA3546"))  # Match fill colours to point colours

# Display the plot
NMDS_Geology

# Save the plot as a .png file
ggsave("NMDS_Geology.png", plot = NMDS_Geology, width = 8, height = 6, dpi = 300)
```

```{r}
#STEP 1.2.6: Plot NMDS by elevation group
# Convert elevation_group to a factor with specified levels and labels
Records_data$elevation_group <- factor(Records_data$elevation_group,                                     
                                       levels = c("1", "2", "3", "4", "5", "6"), 
                                       labels = c("0-199m", "200-399m", "400-599m", "600-799m", "800-999m", "1000-1200m"))

# Align elevation_group with z.points
z.points <- z.points %>%
    mutate(elevation_group = Records_data$elevation_group, 
           forest_type = Records_data$forest_type)

# Check the data for completeness
print(nrow(z.points))
print(unique(z.points$elevation_group))

# Preprocess z.points to remove NAs and group by elevation_group
z.points_clean <- z.points %>%
    filter(!is.na(elevation_group))  # Exclude NAs in elevation_group

# Confirm the cleaned data is not empty
print(nrow(z.points_clean))
print(unique(z.points_clean$elevation_group))

# Calculate convex hulls for each elevation_group
convex_hulls <- z.points_clean %>%
    group_by(elevation_group) %>%
    filter(n() >= 3) %>%  # Ensure there are at least 3 points per group for convex hull
    slice(chull(MDS1, MDS2)) %>%
    ungroup()

# Check convex hulls
print(convex_hulls)

# Plot NMDS by elevation group with convex hulls
NMDS_Altitude <- ggplot(z.points_clean, aes(x = MDS1, y = MDS2)) + 
    # Add convex hull polygons
    geom_polygon(data = convex_hulls, aes(x = MDS1, y = MDS2, fill = elevation_group, group = elevation_group), 
                 alpha = 0.3, colour = "black") +
    # Add points
    geom_point(size = 3, aes(shape = forest_type, colour = elevation_group)) + 
    # Custom theme
    theme(
        axis.text.y = element_text(colour = "black", size = 10, face = "bold"), 
        axis.text.x = element_text(face = "bold", size = 10, colour = "black"), 
        legend.text = element_text(size = 10, face = "bold", colour = "black"), 
        legend.position = "right", 
        axis.title.y = element_text(face = "bold", size = 10), 
        axis.title.x = element_text(face = "bold", size = 10, colour = "black"), 
        legend.title = element_text(size = 10, colour = "black", face = "bold"), 
        panel.background = element_blank(), 
        panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
        legend.key = element_blank()
    ) + 
    # Labels and colours
    labs(x = "NMDS1", colour = "Elevation Group", y = "NMDS2", shape = "Forest Type", fill = "Elevation Group") + 
    scale_colour_manual(values = c("#9EA93F", "#60712F", "#68A357", "#315C2B", "#274029", "#090809")) +
    scale_fill_manual(values = c("#9EA93F", "#60712F", "#68A357", "#315C2B", "#274029", "#090809"))  # Match fill colours to point colours

# Display the plot
print(NMDS_Altitude)

# Save the plot as a .png file
ggsave("NMDS_Altitude.png", plot = NMDS_Altitude, width = 8, height = 6, dpi = 300)
```

################################################################### 
**AIM 2: Descriptive statistics, Diversity with altitude**
###################################################################

```{r}
#STEP 2.1.1: Join the two dataframes based on the site_ID column to add the elevation column
df_reshaped <- df_reshaped %>%
  left_join(Metadata %>% select(site_ID, elevation), by = "site_ID")

# Create the scatterplot with a linear model, formula annotation, and custom scales
species_richness_plot <- ggplot(df_reshaped, aes(x = elevation, y = Count)) +
  geom_point() +  # Add points to the plot
  geom_smooth(method = "lm", fill = "lightblue", color = "darkblue", alpha = 0.5) +  # Add a linear model with light blue shading and dark blue line
  stat_regline_equation(
    aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
    label.x = Inf, label.y = Inf, hjust = 1, vjust = 1
  ) +  # Add the linear model formula and R² value
  scale_x_continuous(limits = c(0, 1100)) +  # Set x-axis scale
  scale_y_continuous(limits = c(0, 55)) +  # Set y-axis scale
  labs(title = "Species Richness by Elevation",
       x = "Elevation (meters)",
       y = "Species Richness") +
  theme_minimal()  # Apply a minimal theme for a clean look

# Save the plot as a PNG file
ggsave("species_richness_by_elevation.png", plot = species_richness_plot, width = 8, height = 6, dpi = 300)
```

END DOCUMENT
