---
title: "Ch3_ClusterAnalysis"
author: "Tessa R Smith"
date: "2024-03-01"
output: html_document
---

## INTRODUCTION

**BEETLE BIOGEOGRAPHY PROJECT**, University of Tasmania. Data collected by Tessa R Smith and volunteers (2020-2022). Code by Tessa Smith, last edited December 2024.

**AIMS**

-   **AIM 4:Exploratory data analysis, Cluster analysis by site**

## DATA PREPARATION ##

```{r, echo = FALSE}
#STEP 0.1: Setup packages and working directory
rm(list = ls()) # clear memory and start afresh

#Change the location for your computer. 
#NOTE make sure you use \\ instead of \ or /, as these result in errors. 
setwd("C:\\Users\\Tessas\\OneDrive - University of Tasmania\\Tessa_PhD\\WORK_Code\\CH3_BROADSCALE\\CH3_ClusterAnalysis_20241114")
```
Notes: works

```{r}
#Load packages
library(dplyr)
library(tidyverse)
```

```{r, echo = FALSE}
#STEP 0.2: Import data
#Metadata (includes altitude etc)
#Reference: Own data (Tessa Smith PhD)
#File type: CSV
#Format: long form. #Columns: site_ID, species, specimen_count. #Rows: Individual records
beetle_records.df <- read.csv("TS_Site_BeetleRecords_20241212.csv", header = TRUE, stringsAsFactors = FALSE)
```
Notes: works

Have removed:
-missing values
-moss sites

```{r, echo = FALSE}
# STEP 0.2.1: Remove rows with no value in the species column
beetle_records_cleaned <- beetle_records.df %>%
  filter(!is.na(species) & species != "")

# STEP 0.2.2: Reshape species list of beetle records from long form into site.count form (wide)
# Only necessary once before start of analysis
df_reshaped <- beetle_records_cleaned %>%
  # Change from long to wide format
  pivot_wider(
    id_cols = site_ID,
    names_from = species,
    values_from = species_count,
    # Sum the entries where there is more than 1 value in specimen_count
    values_fn = list(species_count = sum)
  )  

# STEP 0.2.4: Change NAs in the new dataset to 0
df_reshaped[is.na(df_reshaped)] <- 0

# STEP 0.2.5: Remove any rows with NA
df_reshaped <- na.omit(df_reshaped)

# View the reshaped data
view(df_reshaped)
```
Notes: works

```{r}
#STEP 0.2.5: Log transform data
for (i in 2:675) {
  df_reshaped[, i] = log(df_reshaped[, i] + 1)
}
head(df_reshaped)
```
Notes: 

```{r}
#STEP 0.2.6: Add new columns, Count abundance number for each site, Count species number for each site, Count unique species for each site
df_reshaped_counts <- df_reshaped %>% 
  mutate(abundance = rowSums(.[, -1]),  # Sum all values from columns other than the first one
    species_count = rowSums(.[-1] > 0),  # Count numbers of columns >0
    unique_species = rowSums((.[, -1] > 0) * (colSums(.[-1] > 0) == 1))  # Count number of columns with a value >0 that occur in no other row
  )
view (df_reshaped_counts)
```
Notes: works

```{r, echo = FALSE}
#STEP 0.3: Import data (environmental data(local))
#Metadata (includes altitude etc)
#Reference: Own data (Tessa Smith PhD)
#File type: CSV
#Format: wide form #Columns: environmental variables #Rows:Sites
Metadata <- read.csv("TS_Site_Enviro_metadata_20230622.csv", header = TRUE, stringsAsFactors = FALSE)
View(Metadata) #Opens data in another tab so it can be easily viewed

```
Notes: works

IBRA Bioregion colours
TSR01-#fde725
TSE01-#a0da39
BEL01- #4ac16d
FUR01- #1fa187
TNS01-#277f8e
KIN01- #365c8d
TWE01-#46327e
TCH01-#440154

**AIM 4:Exploratory data analysis, Cluster analysis by site**
Aim: Do sites cluster together?

Code from: https://www.datanovia.com/en/blog/cluster-analysis-in-r-simplified-and-enhanced/
```{r}
#STEP 4.1.1: Load packages
library(factoextra)
library(ggplot2)
```

```{r}
#STEP 4.1.2: Store the site names (first column) in a separate vector
site_names <- df_reshaped[, 1]

#STEP 4.1.3: Convert species occurrences to numeric values
df_reshaped[, 2:688] <- lapply(df_reshaped[, 2:688], function(x) as.numeric(as.character(x)))
```
Notes: works

```{r}
#STEP 4.1.3: Remove sites 77, 101, 124
df_reshaped <- df_reshaped[-c(77, 101, 124), ]
df_reshaped <- df_reshaped[-c(111, 132), ]

```
Notes: these were causing issues with the k-means clustering (very variable samples)

```{r}
# STEP 4.2.1: Correlation-based distance method (Pearson's correlation)
# Exclude the first column (site names) when calculating the distance
res.dist <- get_dist(df_reshaped[, -1], method = "pearson")

# Extract site names from the first column of df_reshaped
site_labels <- as.character(df_reshaped$site_ID)  # Assuming the first column is named "site_ID"

# Convert the distance matrix into a data frame for analysis
dist_matrix <- as.matrix(res.dist)

# Get the upper triangular values (excluding diagonal) for comparison
dist_values <- dist_matrix[upper.tri(dist_matrix, diag = FALSE)]

# Extract row and column indices of the upper triangular matrix
site_combinations <- which(upper.tri(dist_matrix, diag = FALSE), arr.ind = TRUE)
site_names_pairs <- data.frame(
  Site1 = site_labels[site_combinations[, 1]],
  Site2 = site_labels[site_combinations[, 2]],
  Correlation = dist_values
)

# Sort by highest correlations and select the top 20
top_20_correlations <- site_names_pairs %>%
  arrange(desc(Correlation)) %>%
  head(20)

# Print the top 20 pairs with the highest correlations
cat("Top 20 pairs of sites with the highest correlations:\n")
print(top_20_correlations)

# Classify distances >0.75 and <=0.75
num_greater_75 <- sum(dist_values > 0.75)
num_less_equal_75 <- sum(dist_values <= 0.75)

# Print the results
cat("Number of distances > 0.75:", num_greater_75, "\n")
cat("Number of distances <= 0.75:", num_less_equal_75, "\n")

# Transform the distance matrix back to correlation values
correlation_matrix <- 1 - dist_matrix  # Assuming the transformation was 1 - r

# Visualize the correlation matrix with fviz_dist
p <- fviz_dist(as.dist(1 - correlation_matrix), lab = site_labels, 
               gradient = list(low = "blue", mid = "white", high = "red")) +
     theme(axis.text.y = element_text(size = 4, angle = 0, face = "bold"),
           axis.text.x = element_text(size = 4, angle = 90, hjust = 1, vjust = 0.5, face = "bold")) +
     labs(x = "Sites", y = "Sites")

# Print the plot
print(p)
```
Notes: The correlation-based distance is defined by subtracting the correlation coefficient from 1.

Other options for correlation distances are:
Eisen Cosine Correlation Distance (A special case of Pearson’s correlation with means replaced by zero)
Spearman Correlation Distance (Computes the correlation between the ranks of variables)
Kendall Correlation Distance (Measures correspondence between rankings)

```{r}
# Remove columns with constant values
df_reshaped_clean <- df_reshaped %>%
  select_if(function(col) length(unique(col)) > 1)

# STEP 4.4: Enhanced k-means clustering
# Exclude the first column (site names) when performing clustering
res.km <- eclust(df_reshaped_clean[, -1], "kmeans", nstart = 25, k.max = 10, graph = FALSE)

# Convert the site names to a character vector
site_labels <- as.character(df_reshaped_clean[, 1])

# Visualize the clustering result
p <- fviz_cluster(res.km, geom = "point", labelsize = 4) + 
     theme(axis.text.y = element_text(size = 4),
           axis.text.x = element_text(size = 4, angle = 90, hjust = 1, vjust = 0.5))

# Print the plot
print(p)
```
Notes:works.
Need to remove sites 77, 101, 124

```{r}
# STEP 4.5: Gap statistic plot
if (!is.null(res.km$gap_stat)) {
  fviz_gap_stat(res.km$gap_stat)
} else {
  cat("Gap statistic data not available in res.km.\n")
}

# STEP 4.6: Silhouette plot
if (!is.null(res.km$silinfo)) {
  fviz_silhouette(res.km)
} else {
  cat("Silhouette data not available in res.km.\n")
}

# STEP 4.7.1: Optimal number of clusters using gap statistics
if (!is.null(res.km$gap_stat)) {
  cat("Optimal number of clusters using gap statistics:", res.km$nbclust, "\n")
} else {
  cat("Gap statistic data not available for determining optimal clusters.\n")
}

# STEP 4.7.2: Print clustering results
print(res.km)
```
Notes: works 

```{r}
# STEP 4.8: Enhanced hierarchical clustering
# Exclude the first column (site names) when performing clustering
res.hc <- eclust(df_reshaped_clean[, -1], "hclust", scale = "none") # compute hclust

# Create the dendrogram with customized text size and rotation for labels
dendro <- fviz_dend(res.hc, rect = TRUE, show_labels = TRUE, labelsize = 4) + 
  theme(axis.text.x = element_text(size = 4, angle = 90, hjust = 1, vjust = 0.5),
        axis.text.y = element_text(size = 4))

# Print the dendrogram
print(dendro)
```

```{r}
fviz_silhouette(res.hc) # silhouette plot

fviz_cluster(res.hc) # scatter plot
```
