---
title: "TS_CH3_Code"
author: "Tessa R Smith"
date: "2023-09-11"
output: word_document
---

## INTRODUCTION

**BEETLE BIOGEOGRAPHY PROJECT**, University of Tasmania. Data collected by Tessa R Smith and volunteers (2020-2022). Code by Tessa Smith, last edited June 2024.

**AIMS**
-   **AIM 1:Exploratory data analysis, Commute Abundance + Alpha diversity**
-   **AIM 2:Exploratory data analysis, Commute Gamma diversity**
-   **AIM 3:Exploratory data analysis, Commute Beta diversity**
-   **AIM 4:Exploratory data analysis, Commute Zeta diversity**

## DATA PREPARATION

```{r, echo = FALSE}
#STEP 0.1.1: Setup packages and working directory
rm(list = ls()) # clear memory and start afresh

#STEP 0.1.2:Change the location for your computer. 
#NOTE make sure you use \\ instead of \ or /, as these result in errors. 
setwd("C:\\Users\\Tessas\\OneDrive - University of Tasmania\\Tessa_PhD\\WORK_Code\\CH3_BROADSCALE\\CH3_AIM1_DiversityIndices")
```
Notes: works

```{r}
#STEP 0.1.3: Load packages
library(dplyr)
library(tidyverse)
```
Notes: works

```{r, echo = FALSE}
#STEP 0.2.1: Import data
#Metadata (includes altitude etc)
#Reference: Own data (Tessa Smith PhD)
#File type: CSV
#Format: long form. #Columns: site_ID, species, specimen_count. #Rows: Individual records
beetle_records.df <- read.csv("TS_Site_BeetleRecords_20240306.csv", header = TRUE, stringsAsFactors = FALSE)
View(beetle_records.df) #Opens data in another tab so it can be easily viewed
```
Notes: works

```{r, echo = FALSE}
#STEP 0.2.2: Reshapes species list of beetle records from long form into site.count form (wide)
#Only necessary once before start of analysis
df_reshaped <- beetle_records.df %>%
#Change from long to wide format
pivot_wider(id_cols = site_ID,
            names_from = species,
            values_from = species_count,
            # Sum the entries where there is more than 1 value in specimen_count
            values_fn = list(species_count= sum)) 

#STEP 0.2.3: Change NAs in the new dataset to 0
df_reshaped[is.na(df_reshaped)] <-0 
df_reshaped <- na.omit(df_reshaped)

#STEP 0.2.4: Add new columns, Count abundance number for each site, Cound species number for each site, Count unique species for each site
df_reshaped <- df_reshaped %>% 
  mutate(
    abundance = rowSums(.[, -1]),  # Sum all values from columns other than the first one
    species_count = rowSums(.[-1] > 0),  # Count numbers of columns >0
    unique_species = rowSums((.[, -1] > 0) * (colSums(.[-1] > 0) == 1))  # Count number of columns with a value >0 that occur in no other row
  )
view (df_reshaped)
```
Notes: works

```{r}
#STEP 0.2.5: Log transform data
for (i in 2:688) {
  df_reshaped[, i] = log(df_reshaped[, i] + 1)
}
head(df_reshaped)
```
Notes: works. Logarithmic transformation (log(x + 1)) is frequently used to stabilize variance and normalize the distribution of species abundance data. It is particularly useful when dealing with highly skewed data, as it compresses the range of values and reduces the impact of large outliers. 

```{r}
#STEP 0.2.6: Remove rare species (species present in <5 sites)

# Function to remove columns where species is present in less than 5 sites
remove_infrequent_species <- function(df, threshold = 5) {
  # Calculate the number of sites each species is present in
  species_counts <- colSums(df != 0)
  
  # Identify columns to keep and to remove
  columns_to_keep <- species_counts >= threshold
  columns_to_remove <- species_counts < threshold
  
  # Filter the data frame
  df_filtered <- df[, columns_to_keep]
  
  # Print the number of columns included and removed
  num_included <- sum(columns_to_keep)
  num_removed <- sum(columns_to_remove)
  cat("Number of columns included:", num_included, "\n")
  cat("Number of columns removed:", num_removed, "\n")
  
  return(df_filtered)
}

# Apply the function to df_reshaped
df_filtered <- remove_infrequent_species(df_reshaped, 5)
```
Notes: works. Removes the majority of the data though. 

```{r, echo = FALSE}
#STEP 0.3.1: Import data (environmental data(local))
#Metadata (includes altitude etc)
#Reference: Own data (Tessa Smith PhD)
#File type: CSV
#Format: wide form #Columns: environmental variables #Rows:Sites
Metadata <- read.csv("TS_Site_Enviro_metadata_20230622.csv", header = TRUE, stringsAsFactors = FALSE)
View(Metadata) #Opens data in another tab so it can be easily viewed

```
Notes: works

## DATA ANALYSIS

-   **AIM 1:Exploratory data analysis, Commute Abundance + Alpha diversity**
Method reference: Numerical Ecology with R book
(<https://rpubs.com/CPEL/NMDS>)

```{r}
#STEP 1.1: Load packages
library(vegan)
library(ggplot2)
library(ggpmisc)
```
Notes: works

```{r}
#STEP 1.2.1: Plot diversity vs. abundance by site

# Fit the linear model
lm_fit <- lm(species_count ~ abundance, data = df_reshaped)

# Extract the coefficients
intercept <- coef(lm_fit)[1]
slope <- coef(lm_fit)[2]

# Calculate R-squared
r_squared <- summary(lm_fit)$r.squared

# Extract the p-value
p_value <- summary(lm_fit)$coefficients[2, 4]

# Create formula text
formula_text <- paste0("y = ", round(intercept, 2), " + ", round(slope, 2), "x",
                       "\nR² = ", round(r_squared, 2),
                       "\np = ", format.pval(p_value, digits = 2))

# Make the scatterplot with a linear model and formula
ggplot(df_reshaped, aes(x = abundance, y = species_count)) +
  geom_point() +
  geom_smooth(method = "lm", col = "blue", se = FALSE) +
  labs(title = "Scatterplot of Species Counts vs Abundance with Linear Model",
       x = "Abundance",
       y = "Species Diversity") +
  theme_minimal() +
  annotate("text", x = Inf, y = Inf, label = formula_text, hjust = 1.1, vjust = 1.1, size = 5)
```
Notes: works

```{r}
#STEP 1.2.2: Select the top 5 rows based on the abundance column
top_5_rows <- df_reshaped %>%
  arrange(desc(abundance)) %>%
  head(5)

# Print the top 5 rows and their abundance values
print(top_5_rows)

# Print only the abundance values of the top 5 rows
print(top_5_rows$abundance)
```
Notes: works

```{r}
#STEP 1.2.3: Ensure all columns are numeric, excluding the last three columns
df_numeric <- df_reshaped %>% select_if(is.numeric) %>% select(-((ncol(.) - 2):ncol(.)))

# Calculate the total abundance for each species
species_abundance <- colSums(df_numeric)

# Convert to a data frame
species_abundance_df <- data.frame(species = names(species_abundance), abundance = species_abundance)

# Select the top 5 species based on abundance
top_5_species <- species_abundance_df %>%
  arrange(desc(abundance)) %>%
  head(5)

# Print the top 5 species and their abundance values
print(top_5_species)
```
Notes:works

```{r}
#STEP 1.3.1: Assuming df_reshaped_counts is your dataframe
# Calculate statistics
min_species_counts <- min(df_reshaped$species_count, na.rm = TRUE)
max_species_counts <- max(df_reshaped$species_count, na.rm = TRUE)
mean_species_counts <- mean(df_reshaped$species_count, na.rm = TRUE)
sd_species_counts <- sd(df_reshaped$species_count, na.rm = TRUE)

# Print the statistics
cat("Minimum species counts:", min_species_counts, "\n")
cat("Maximum species counts:", max_species_counts, "\n")
cat("Mean species counts:", mean_species_counts, "\n")
cat("Standard deviation of species counts:", sd_species_counts, "\n")
```
Notes: works

```{r}
#STEP 1.3.2: Select the top 5 rows based on the species richness column
top_5_rows_richness <- df_reshaped %>%
  arrange(desc(species_count)) %>%
  head(5)

# Print the top 5 rows and their abundance values
print(top_5_rows_richness)

# Print only the abundance values of the top 5 rows
print(top_5_rows_richness$species_count)
```
Notes: works

```{r}
#STEP 1.4.1: Assuming df_reshaped_counts is your dataframe
# Calculate statistics
min_species_unique <- min(df_reshaped$unique_species, na.rm = TRUE)
max_species_uniqu <- max(df_reshaped$unique_species, na.rm = TRUE)
mean_species_uniqu <- mean(df_reshaped$unique_species, na.rm = TRUE)
sd_species_uniqu <- sd(df_reshaped$unique_species, na.rm = TRUE)

# Print the statistics
cat("Minimum species counts:", min_species_counts, "\n")
cat("Maximum species counts:", max_species_counts, "\n")
cat("Mean species counts:", mean_species_counts, "\n")
cat("Standard deviation of species counts:", sd_species_counts, "\n")
```
Notes: works

```{r}
#STEP 1.4.2: Select the top 5 rows based on the unique species column
top_5_rows_unique <- df_reshaped %>%
  arrange(desc(unique_species)) %>%
  head(5)

# Print the top 5 rows and their abundance values
print(top_5_rows_unique)

# Print only the abundance values of the top 5 rows
print(top_5_rows_unique$unique_species)
```
Notes: works

-   **AIM 2:Exploratory data analysis, Commute Gamma diversity**
```{r}
#STEP 2.1: Load packages
library(vegan)
```
Notes: works

```{r}
#STEP 2.2: Generate a species accumulation curve

# Ensure all columns are numeric
df_numeric <- df_reshaped %>% select_if(is.numeric)

# Create a species accumulation curve
spec_accum <- specaccum(df_numeric, method = "random")

# Plot the species accumulation curve
plot(spec_accum, 
     xlab = "Number of Sites", 
     ylab = "Accumulated Number of Species", 
     main = "Species Accumulation Curve", 
     ci.type = "polygon", 
     col = "blue", 
     lwd = 2, 
     ci.lty = 0, 
     ci.col = "lightblue")

```
Notes: works

-   **AIM 3:Exploratory data analysis, Commute Beta diversity**

```{r}
#STEP 3.1: Load packages
library(vegan)
library(dplyr)
```
Notes: works

```{r}
#STEP 3.2.1: Plots betadiversity  with vegan. Created triangular plot based on pairwise comparisons between assemblages. 
m <- betadiver(df_reshaped[, -1])
plot(m)
```
Notes: works

```{r}
#STEP 3.2.2: Extract names of highly correlated pairs
```
Notes: not working

```{r}
#STEP 3.3.1: Compute diversity indices

# Extract site names from the first column
site_names <- df_reshaped[, 1]

# Ensure all columns are numeric, excluding the site names and the last three columns
df_numeric <- df_reshaped %>% select_if(is.numeric) %>% select(-((ncol(.) - 2):ncol(.)))

# Compute diversity indices  
N0  <- rowSums(df_numeric > 0) # Species richness
H   <- diversity(df_numeric) # Shannon entropy
N1  <- exp(H) # Shannon diversity number
N2  <- diversity(df_numeric, "inv") # Simpson diversity number
J   <- H / log(N0) # Pielou evenness
E1  <- N1 / N0 # Shannon evenness (Hill's ratio)
E2  <- N2 / N0 # Simpson evenness (Hill's ratio)

# Combine diversity indices into a data frame with site names
div <- data.frame(Site = site_names, N0, H, N1, N2, E1, E2, J)

# Print the diversity indices table
print(div)

# Export the div dataframe to a CSV file
write.csv(div, "diversity_indices.csv", row.names = FALSE)
```
Notes: works

-   **AIM 4:Exploratory data analysis, Commute Zeta diversity**
```{r}
#STEP 4.1: Load packages
library(zetadiv)
```
Notes: works

```{r}

#STEP 4.2.1: Remove the first column if it contains site names and the last three columns
df_numeric <- df_reshaped[, -c(1, (ncol(df_reshaped) - 2):ncol(df_reshaped))]

# Ensure df_numeric is in the correct format (species in columns, sites in rows)
# Convert to presence/absence data if necessary
df_pa <- df_numeric > 0

# Define the number of sites to consider in the zeta diversity calculation
# For example, calculate zeta diversity for up to 5 sites
zeta_max_order <- 5

# Calculate zeta diversity
zeta_values <- Zeta.decline.ex(df_pa, orders = 1:zeta_max_order)

# Print the zeta diversity values
print(zeta_values)

# Optionally, plot the zeta diversity values
plot(zeta_values$zeta.val, type = "b", xlab = "Order", ylab = "Zeta Diversity", main = "Zeta Diversity Plot")
```
Notes: works

END DOCUMENT
